Search for Spatially Extended Fermi-LAT Sources Using Two Years of Flight Data

### 2011-07-08 extended_catalog/paper/extended_catalog.tex
adding green catalog

% TODO:
%  how to format powerlaw, power-law, power law?

@@

% some units
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_\text{ext}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_\text{inc}}\xspace}

@@

% * Resolving spatially extended Fermi sources is important for 
%    * studying supernova remnants, 
%    * pulsar wind nebulae, 
%    * galaxies,  
%    * potentially dark matter or new physics.
% * We searched all 2FGL catalog sources
% * Systematic test for a certain kind of (~ small) regularly shaped extended

@@

Morphological studies of sources in the \gev energy range
using the Fermi LAT are challenging because of the significantly energy
dependent PSF and because of systematic errors in the galactic diffuse
emission.  The LAT's PSF is determined at lower energies by multiple
scattering in the tracker and the 68\% containment radius of the PSF is
approximately 5.1\deg at 100 \mev  when averaged over instrument acceptance
and including photons which convert in either the thick or thin layers of
the tracker. The PSF approaches a limit given by the granularity of the
tracker channels and is 0.14\deg at 10 \gev (\cite{on_orbit_calibration}).
Therefore, the higher energy photons are significantly more important for
resolving a source but are limited by statistics. Therefore,
a detailed analysis procedure mused be used to maximize our
sensitivity.

@@

A new analysis tool has been developed to address the unique requirements
for studying spatially extended sources with the Fermi LAT. The tool
provides a maximum likelihood analysis where the Poisson likelihood
for observing the measured counts is computed is maximized by fitting
parameters of the model.  The extension of a source can be modeled by a
geometric shape (e.g. a disk or Gaussian) and the source's position
and extension can simultaneously fit by maximizing the likelihood.

@@

\tsinc can not be directly compared to \tsext to see whether two
point sources are preferred over one extended source. Unfortunately,
performing a likelihood ratio test between two models that are not nested
(\cite{statistics_with_care}). So the comparison of \tsext with \tsinc is
not a calibrated test.  Nevertheless, we find the extreme cases where
$\tsinc \ll \tsext$ or $\tsinc\gg\tsext$ to be suggestive and use this
as a test for source confusion. For the spatially extended candidates
found, we computed both \tsext and \tsinc.  The results are presented
in table~\ref{dual_localization_results}

@@

\todo{This Section Needs Work}

@@

This is demonstrated in figure~\ref{compare_disk_gauss}.  It shows the
PSF that would be observed for a powerlaw source of spectral index two,
the probability density function for a Gaussian and uniform disk of width
$0.5\deg$, and the probability density function for the same Gaussian
and uniform disk convolved with the PSF.  The figure shows the analysis
using an energy range between 1 \gev and 100 \gev and using an energy
range between 10 \gev and 100 \gev.  The convolved shape of the disk and
Gaussian do separate better at higher energy where the PSF is narrower.
This plot shows that the LAT is not very sensitive to probing different
kinds of extended sources.  In this paper we always fit with uniform disk
spatial models.

@@

Finally, we note that comparing fit sizes to counts maps can look
misleading.  This effect can best be seen in figure~\ref{compare_r68}
which shows a simulated a extended source with a uniform disk spatial
model with $\sigma=0.5\deg$. The simulated source has integral flux
between 1 \gev and 100 \gev of $3\times 10^{-8}\ph/\cm^2/s$, and a
spectral index of 2.  To reduce Poisson noise, figure~\ref{compare_r68}
shows a counts map of this source smoothed with by a 0.1\deg Gaussian
kernel. Overlaid on the plot is the results of fitting this simulated
source with a uniform disk and a Gaussian spatial model.  The solid white
line represents the best fit disk extension $\sigma_\text{disk}=0.496$
The solid black line represents the fit extension $\sigma$ for the
Gaussian which was found to be $\sigma_\text{Gaussian}=0.2627$.
The white and black dashed lines are the 68\% containment radius of
the fit Disk and Gaussian respectivly.  Even though the fit $\sigma$
for the disk and uniform Gaussian seem very different, the fit 68\%
containment radius are very similar ($\rsixeight_\text{,disk}=0.41$
whereas $\rsixeight_\text{,Gaussian}=0.40$) and the results are in
agreement.

This study shows that depending upon what spatial model you pick and
what parameter you choose to overlay on the smoothed counts ($\sigma$ or
$rsixeight$), you will get shapes that look very different but predict
very similar counts distributions.  Since we are not very sensitive to
the particular morphology of an extended source, there is no correct
choice for which shape to use. For the following analysis, we always use
a uniform disk as our spatial model because of its conceptual simplicity
and to emphasize our ignorance to the true morphology. And for subsequent
smoothed counts maps, we overlay the $\sigma$ because that shows directly
the edge of the shape we fit.  But due to smearing of the PSF, it has the
unfortunate consequence that the extended source shape always (including
in the simulated image) looks larger than in the counts map.

The second conclusion we make is that because a counts map is smeared
out by the PSF and depends strongly on the particular color bar used, it
is very difficult to determine directly from a counts map a reasonable
spatial model that fits the data.  It is much better to obtain the
spatial model by a direct fit of a spatial parameters of a selected
model to the data.  The third conclusion we make is that when comparing
fit extensions to those at other wavelenghts (especially in \tev), it is
important to compare fits of consistent spatial models. If this is not
possible, it is important to at least compare indirectly the $rsixeight$
for the different spatial models.

@@

On the other hand, since the null hypothesis rests on the edge of
parameter space when the extension is 0, Wilks' Theorem does not hold
(\cite{warn_wilks_theorem}).  One might expect the actual distribution
to be a half $\chi^2_1$ distribution with one degree of freedom plus
half a delta function at 0:
\begin{equation}
  P(\ts)=\tfrac{1}{2}(\chi^2_1(\ts)+\delta(\ts)).
\end{equation}
Chernoff proved that under certain assumptions, this is the distribution
that would hold when the null hypothesis lies on the edge of parameter
space (\cite{chernof}).  The false detection
probability can be estimated by integrating this function from an observed
test statistic value to infinity. It is for this particular distribution
that the often quoted result holds that $\sqrt{\ts}$ is a measure of
the number of $\sigma$ of the detection (\cite{mattox_egret}).

@@

The Monte Carlo study involved simulating point sources of varying
spectral models and fitting them using \pointlike as both point
and extended sources.  Fermi LAT data was simulated using program
\gtobssim\footnote{\gtobssim is distributed by and supported by the
Fermi Science Support Center (\cite{fssc}).}  The sources were simulated
with a powerlaw spectral model with 6 Were given 6 $>100$ \mev fluxes
ranging from $3\times 10^{-9} \ph/\cm^2/s$ to $10^{-6} \ph/\cm^2/s$.
The sources were then simulated with spectral index ($\gamma$) values of
1.5, 2, 2.5, and 3.  These values were picked to probe a representative
range measurable values.  For particular spectral indices, only the fluxes
which would create a statistically significant source detection ($\ts>25$)
where selected, since we are only interested in testing for extension
sources with a significant extent. The sources were simulated on top of
an isotropic powerlaw background with a Sreekumar-like spectrum ($>100$
\mev flux of 1.5e-5 and spectral index 2.1) (\cite{sreekumar_isotropic}).

@@

Our Monte Carlo study showed broad agreement between simulations and
Wilks' theorem and that we can use \tsext as a test of the statistical
significance of a source's extension.  Nevertheless, the agreement
is not perfect.  It should be noted that the discrepancy seems to be
worst for bright sources. This seems to imply that numerical errors in
the convolution become more apparent for sufficiently bright sources.
Other possible reasons for departure from Wilks' theorem might include
\pointlike ignoring energy dispersion. Energy dispersion will change
the shape as a function of energy. \pointlike may also run into possibly
problems with the fitting algorithm getting stuck in false minimum. It
is important to emphasize that if our emperical curve lies to the left of
the curve we assume to estimate significance, we would underestimate the
statistical significance of our result. Since this is the case for most
of our simluations, we are reasonable confident that $\sqrt{\tsext}$
can be safely used as a measure of the statistical significance of a
source's extension.

@@

We performed another Monte Carlo study to determine the Fermi LAT's
sensitivity to spatially extended sources. The sensitivity to extension
is defined as the flux at which 
the average value of $\tsext$ for a source is 
e equal to 25 (for a given spectral model). This correseoncs
to an average extension significance of $5\sigma$.  This is not the
sensitivity to detect an extended source but instead the sensitivity to
resolve an already detected source as being extended.

One would intuitivly expect that we would be less sensitivie to resolving
small extended sources since they would be lost in the PSF.  Similarly,
we would be less sensitive to very large extended sources which would
get completely lost in the background. Furtheremore, one expects our
sensitivity to be greater for harder sources since our angular resolution
improves significantly with energy.

To calculate the LAT's sensitivity to spatially extended sources,
we simulated sources with a uniform radially symmetric
disk morphology on top of a powerlaw Sreekumar-like isotropic
spectrum (\cite{sreekumar_isotropic}).  The Monte Carlo
simulation was perforormed over a one year time interval using a default
rocking profile and an assumed a represntative livetime fraction of 0.8.
For a given extension and spectral index, we calculated the
sensitivity by varying the flux of a source until we found 
when $\tsext=25$. In particular, we hand picked a flux range
where $\tsext<25$ and $\tsext>25$ and simulated 10 points with
a log uniform distribution in this range and performed an extension
test for each of these fluxes. Since the measured signifciance is
a statistical which will vary from realization to realization,
we then repeated this procedure $>15$ times for each spectral
and spatial model.  By fitting a line to the flux vs $\tsext$, we were
able to estimage the flux at which $\tsext=25$.

Our first result is the extension sensitivity for sources of varying
spectral indices. The result is shown in figure~\ref{index_sensitivity}.
We computed the sensitivty for sources with a spectral index of 1.5,
2, 2.5, and 3 and for sources of 20 extensions varying from $0.1\deg$
to $2.0\deg$. The $y$ axis is the 100 \mev to 100 \gev flux of the
source and the $x$ axis is the Monte Carlo extension. The curves then
represent the division in this extension and flux space where one gets
on average $\tsext=25$.
The solid lines of the plot are the sensitivity computed when using
with all photons with energry between 100 \mev and 100 \gev. The dashed lines is the
same sensitivity but only using photons with energy from 1 \gev to 100 \gev.

This results shows that our extension sensitivity is
significantly better for harder sources where there are more high
energy photons. Furthermore, our sensitivity is best for sources with an
extension around $0.5\deg$.  our sensitivity gets significantly worse
for very small sources and gets slightly worse for very large sources.

Futheremore, we note that except for the hardest or largest sources,
our sensitivity is almost entirly whether or not we use photons 
with energy from
100 \mev to 1 \gev. On the other hand, other systematic errors become
increasingly important at low energy so for our extension search we will
not use the photons in this energy range.

We performed a second computation to see how the LAT's extension
sensitivity varied with increasing background, as is at lower latitude
in the galactic plane. To do so, we computed the sensitivity using the
extragalactic barground as well as 10 and 100 times the extragalactic
background. The latter result approximatly represents the intensity of
the galactic diffuse emission near the galactic center. The results are
shown in figure~\ref{diff_factor_sensitivity}. Two plots are presented,
one fitting only in the energy range between 1 \gev and 100 \gev and
one fitting only in the energy range between 10 \gev and 100 \gev. For
each plot, the $y$ axis quotes the flux in the fit energy range and the
simulation was done only for a representivite spectral index 2 source.

These plots show that we are much less sensitivie to extended sources in
a higher background region. By comparing the two plots, we also see that
we would be more sensitive to an extended source if it had the same
integral flux, but only over a higher energy range. This is 
consistent with the improved angular resolution at higher energy.

@@

As was shown, we gain very little in sensitivity using photons with energy
below 1 \gev. On the other hand, the large PSF at low energy makes us
more succeptable to systmetic errors arising from source confusion due
to multiple point soruces and succeptable to incorrect modeling of the
galactic diffuse emission.  For that reason, we performed our search
using photons only between 1 \gev and 100 \gev.

We also performed a search for extended sources using photons only between
10 \gev and 100 \gev. Even though we are testing the same sources, this
approach is complimentary because there are regions in the galactic
plane where at lower energies the diffuse emission dominates and is
likely not perfectly modeled. On the other hand, at the highest energies
accessible by the LAT, the galactic diffuse emission will be smaller and
we might expect to detect significantly harder sources. This approach is
also beneficial because harder sources with nearby neighbors that are
softer may end up being incorrectly fit at lower energies but correctly
found looking only at the high energy.  This is also beneficial in the
galactic plane looking for possible extended sources near pulsars which
are very bright at lower energy but cutoff before 10 \gev.  This approach
of using only photons with energy above 10 \gev was succesfully used to detecting
HESS J1825-137 with the Fermi LAT~\cite{fermi_hess_j1825}.

For each catalog source, we modified the spectral index of the source
to be a powerlaw. We did this for two reasons. The first is that for a
source to be significantly extended, we would expect it to be reasonably
hard and therefore be fit reasonably well by a powerlaw.  The seond reason
is that since we begin our analysis at either 1 \gev or 10 \gev, we didn't
expect there to be significant curvature over this smaller energy range.
This method has the advantange that for the following analysis, we can
simply quote the fit powerlaw spectral index.

Finally, when analyzing our region, we automatically removed all other
catalog sources which were within 0.5\deg. This was done because of
a concern that truly extended sources might end up in the catalog as
multiple point sources. These spurious nearby catalogs sources might then
distort our fit enough that it no longer would show up as being extended.
Nevertheless, we investiated for each of our extended candidates if
if we seemed to be legitimatly removing real backgorund sources. And
we employed a dual localiation procedure (described above) to test
all of our candidates for source confusion.

@@

We expect most extended sources to be inside teh galactic plane
since extragalactic sources are typically too far away to be
resolvable. Unfortunately, the \gev emission in the galactic plane
is dominated by anisotropic galactic diffuse emission.  Modeling the
galactic diffuse emission and finding sources on top of this emission is
difficult~\ref{first_diffuse_paper}. This issue has been discussed in the
Fermi catalog~\ref{first_cat,second_cat}. Furthermore, the galactic plane is
very crowded and so it is offent difficult to get a good model for the
region around a catalog source.  Because of this, finding a source with
$\tsext>16$ is not a sufficient test and we perform several crosschecks
to ensure our overal region is well modeled.

As an example of why this issue, we describe our test of P72Y3047 for
extension.  Figure~\ref{example_bad_fit} shows a smoothed counts map
using photons with energy above 1 \gev. In the map, the galacitc and isotropic diffuse
emission subtracted.  A two year catalog source P72Y3047 was found at the
position $(l,b)=(37.79\deg,1.14\deg)$.  By looking at the smoothed map,
we see that there appears to be overall some large scale residaul in the
diffuse emission along the galactic plane. Furthermore, it appears that
we don't have a great model of the nearby catalog sources. As a result,
P72Y3047 fits to a very large extension of 1.37\deg and the result is
statistically significant with \tsext=51.6.  By looking at the residuals,
it is clear that this complicated region is not well fit even though
the source's extension is statistically significant. For this reason,
we have to perform a series of followup analysis for all of the soruces
that are found with \tsext=16.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml

First, we took the spatial and spectral model found using \pointlike
and performed a crosscheck by refitting the spectral parameters using
\gtlike.  We used a square ROI $14.1\deg$ in diamater with $1/16\deg$
pixels and the same energy binning. We found that $1/16\deg$ bins were
fine enough for \gtlike to get an extension significances consistent
with \pointlike.  Using \gtlike, we found a second measure of \tsext and
we only considered a source to be significanlty extended if \pointlike
and \gtlike both agreed.  For the new extended source candidates, with
table~\ref{alt_diff_model_results} shows a comparison of their the value
of \tsext found by \gtlike and \pointlike.  Everywhere else, the values
from \gtlike are used.

Additionaly, we generated for each extended source candidate residual
test statistic (\ts) maps.  The residual \ts map is a map of the \ts
value evalualted at each point by adding onto the map at that position
a new source of spectral index 2 and finding the increase in lieklihood
when fitting its flux. \ts maps are especially useful as looking for
new sources or other residual featuers that are not included in the
model. They also have the advantagous feature that the aboslute scale
can be used to clearly look for when the residual is signifciant.
The downside is they are biased to look for residual features
that look like the PSF and so are not as sensitive to diffuse residual
features. Furthermore, since by defintion \ts>0, \ts maps can only show
positive fluctuations. Nevertheless, \ts maps were useful for
assessing the overal quality of the fit. Figure~\ref{res_tsmap}
shows an exmaple residual \ts map for the extended source IC443.

For each candidate, we also generated counts slices. A counts slice
plots the sum of all counts in the longitudinal direction within 2\deg in
the latitudinal direction.  The same plot is also shown integrating in
the longitudinal direction.  On top of the plot is overlayed the model
predicted counts assuming the source has a disk spatial model, assuming
the source is a point source, and fitting just the diffuse emission in
the region.  An example counts slice is shows for the extended source
IC443 in figure~\ref{counts_slice}. Counts slices were especially useful
for looking at how well the overal diffuse emisison model fits the region
near the sources.

For each candidate, we also made was a radial integral of the data. This
plot shows the sum of all counts within a given distance from the
source's center. The plot is binned uniformly in $\Delta \theta^2$ so
that an isotropic background will be flat in the plot. On top of this
plot is overlayed the model predicted counts assuming the source has a
disk spatial model, assumgng the source is a point source, and fitting
just the diffuse emisison in the region. An example radial integral
is shown for the extended source IC443 in figure~\ref{radial_profile}.
This plot is best for comparing the observed counts to the PSF.

Finally, for each source, we performed the dual localization method
described in section~\ref{dual_localization_method}.  We 
did so by replacing the source with two sources each containing
half of the original flux and then using \pointlike to
simulatnoulsy fit the postion and spectrum of these sources.
We then used \gtlike to refit the spectrum of these
sources using the best fit positions. For the
new extended sources found, we summarize the results of
this dual localization test and compare the values of \tsext
to \tsinc in table~\ref{dual_localization_results}.

@@

Since these sources are all included in the 2 year catalog, we applied
no special method to test these sources but used the same pipeline
that tested all other catalog sources for extension.  The results are
summarized in table~\ref{known_extended_sources}.  This table includes
the best fit position and extension. It also includes the source's
significance \ts and extension significance \tsext and the localization
error ellipse. The table also includes the best fit powerlaw spectral
model.

@@

\subsection{2FGL J1551.3-5333}

>>>>

% TODO:
%  how to format power-law, power-law, power law?

@@

% some units
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\renewcommand{\sec}{\text{s}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_\text{ext}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_\text{inc}}\xspace}

@@

@@

Morphological studies of sources in the \gev energy range
using the LAT are challenging because of the significantly energy
dependent PSF and because of systematic errors in the galactic diffuse
emission.  The LAT's PSF is determined at lower energies by multiple
scattering in the tracker and the 68\% containment radius of the PSF is
approximately 5.1\deg at 100 \mev  when averaged over instrument acceptance
and including photons which convert in either the thick or thin layers of
the tracker. The PSF approaches a limit given by the granularity of the
tracker channels and is 0.14\deg at 10 \gev (\cite{on_orbit_calibration}).
Therefore, the higher energy photons are significantly more important for
resolving a source but are limited by statistics. Therefore,
a detailed analysis procedure mused be used to maximize our
sensitivity.

@@

A new analysis tool has been developed to address the unique requirements
for studying spatially extended sources with the LAT. The tool
provides a maximum likelihood analysis where the Poisson likelihood
for observing the measured counts is computed is maximized by fitting
parameters of the model.  The extension of a source can be modeled by a
geometric shape (e.g. a disk or Gaussian) and the source's position
and extension can simultaneously fit by maximizing the likelihood.

@@

\tsinc can not be directly compared to \tsext to see which model is better
because the models are not nested (\cite{statistics_with_care}). Even
though the comparison of \tsext with \tsinc is not a calibrated test,
we find the extreme cases where $\tsinc \ll \tsext$ or $\tsinc\gg\tsext$
to be suggestive and use this as a test for source confusion. For the
spatially extended candidates found, we computed both \tsext and \tsinc.
The results are presented in table~\ref{dual_localization_results}

@@

\label{compare_source_size}

@@

This is demonstrated in figure~\ref{compare_disk_gauss}.  It shows the
PSF that would be observed for a power-law source of spectral index two,
the probability density function (PDF) for a Gaussian and uniform disk of width
$0.5\deg$, and the PDF for the same Gaussian
and uniform disk convolved with the PSF.  The figure shows the analysis
using an energy range between 1 \gev and 100 \gev and using an energy
range between 10 \gev and 100 \gev.  The convolved shape of the disk and
Gaussian do separate better at higher energy where the PSF is narrower.
This plot shows that the LAT is not very sensitive to probing different
kinds of extended sources.  In this paper we always fit with uniform disk
spatial models.

@@

Overlaying fit sizes on top of counts maps can be misleading.
This effect can best be seen in figure~\ref{compare_r68} which shows
an extended source simulated in the energy range from 1 \gev to 100
\gev with a uniform disk spatial model. The simulated source has a size
$\sigma=0.5\deg$, an integral flux between 1 \gev and 100 \gev of $3\times
10^{-8}\ph/\cm^2/\sec$, and a spectral index 2.  To reduce Poisson noise,
the counts map in figure~\ref{compare_r68} is smoothed by a 0.1\deg
Gaussian kernel. Overlaid on the plot is the fit extension assuming the
source has a uniform disk and a Gaussian spatial model.  The solid white
line represents the best fit disk extension $\sigma_\text{disk}=0.496$
The solid black line represents the best fit extension $\sigma$ for
the Gaussian which was found to be $\sigma_\text{Gaussian}=0.2627$.
The white and black dashed lines are the 68\% containment radius of
the fit Disk and Gaussian respectively.  Even though the fit $\sigma$
for the disk and uniform Gaussian seem very different, their 68\%
containment radius are very similar ($\rsixeight_\text{,disk}=0.41$
and $\rsixeight_\text{,Gaussian}=0.40$).

This study shows that the extension $\sigma$ for different spatial
models can be very different even when predicting almost the same PDF.
Since we are not very sensitive to the particular morphology of an
extended source, there is no correct choice for which shape to plot. For
the following analysis, we use a uniform disk as our spatial model
and overlay the disk's edge $\sigma$.  But due to smearing of the PSF,
$\sigma$ typically looks larger than the observed counts distribution.
When comparing the fit extensions with other wavelengths (especially \tev
measurements of galactic sources), it is important to compare consistent
values. If the assumed shapes are different, it is best to indirectly
compare $rsixeight$.

@@

On the other hand, since the null hypothesis rests on the edge of
parameter space when the extension is 0, Wilks' Theorem does not hold
(\cite{warn_wilks_theorem}).  One might expect the actual distribution
to be a half $\chi^2_1$ distribution with one degree of freedom plus
half a delta function at 0:
\begin{equation}
  P(\ts)=\tfrac{1}{2}(\chi^2_1(\ts)+\delta(\ts)).
\end{equation}
Chernoff proved that under certain assumptions, this is the distribution
that would hold when the null hypothesis lies on the edge of parameter
space (\cite{chernoff}).  The false detection
probability can be estimated by integrating this function from an observed
test statistic value to infinity. It is for this particular distribution
that the often quoted result holds that $\sqrt{\ts}$ is a measure of
the number of $\sigma$ of the detection (\cite{mattox_egret}).

@@

The Monte Carlo study involved simulating point sources of varying
spectral models and fitting them using \pointlike as both point
and extended sources.  LAT data was simulated using program
\gtobssim\footnote{\gtobssim is distributed by and supported by the
Fermi Science Support Center (\cite{fssc}).}  The sources were simulated
with a power-law spectral model with 6 Were given 6 $>100$ \mev fluxes
ranging from $3\times 10^{-9} \ph/\cm^2/\sec$ to $10^{-6} \ph/\cm^2/\sec$.
The sources were then simulated with spectral index ($\gamma$) values of
1.5, 2, 2.5, and 3.  These values were picked to probe a representative
range measurable values.  For particular spectral indices, only the fluxes
which would create a statistically significant source detection ($\ts>25$)
where selected, since we are only interested in testing for extension
sources with a significant extent. The sources were simulated on top of
an isotropic power-law background with a Sreekumar-like spectrum ($>100$
\mev flux of 1.5e-5 and spectral index 2.1) (\cite{sreekumar_isotropic}).

@@

Our Monte Carlo study showed broad agreement between simulations and
Wilks' theorem and that we can use \tsext as a test of the statistical
significance of a source's extension.  Nevertheless, the agreement
is not perfect.  It should be noted that the discrepancy seems to be
worst for bright sources. This seems to imply that numerical errors in
the convolution become more apparent for sufficiently bright sources.
Other possible reasons for departure from Wilks' theorem might include
\pointlike ignoring energy dispersion. Energy dispersion will change
the shape as a function of energy. \pointlike may also run into possibly
problems with the fitting algorithm getting stuck in false minimum. It
is important to emphasize that if our empirical curve lies to the left of
the curve we assume to estimate significance, we would underestimate the
statistical significance of our result. Since this is the case for most
of our simulations, we are reasonable confident that $\sqrt{\tsext}$
can be safely used as a measure of the statistical significance of a
source's extension.

@@

We performed a second study to determine the LAT's sensitivity
to spatially extended sources. The sensitivity to extension
is defined as the flux at which the average value of
$\tsext$ is equal to $\langle\ts\rangle=25$. As was seen in
section~\ref{monte_carlo_validation}, this corresponds to an average
extension significance of $5\sigma$.
Qualitatively, we expect that the LAT would not be sensitive to
extended sources much smaller than the PSF. Similarly,
we expect that the LAT would not be sensitive to sources
much larger than the PSF since they would get lost in the
background.

To calculate the LAT's sensitivity to spatially extended sources,
we simulated sources with a uniform radially symmetric disk
morphology on top of a power-law Sreekumar-like isotropic spectrum
(\cite{sreekumar_isotropic}).  The Monte Carlo simulation was performed
over a one year time interval using a default rocking profile and a
livetime fraction of 0.8.  For a given extension and spectral index, we
first hand picked a flux range such that $\tsext<25$ and $\tsext>25$. We
then picked 10 flux points in this range and performed an extension test
for each of these fluxes. We repeated this procedure $>15$ times for each
spectral and spatial model.  By fitting a line to the flux vs $\tsext$,
we estimated when $\langle\tsext\rangle=25$.

Figure~\ref{index_sensitivity} shows the extension sensitivity for
sources of different spectral indices and extensions.  We selected
spectral indices 1.5, 2, 2.5, and 3 and 
extensions from $\sigma=0.1\deg$ to $2.0\deg$.  The curves
on the plot represent the division in extension and flux space where
$\langle\tsext=25\rangle$.  The plot shows by the sensitivity using all
photons with energy between 100 \mev and 100 \gev and the sensitivity
using only photons with energy between 1 \gev and 100 \gev.  This results
shows that the LAT's sensitivity to extension is better for harder sources
with more high energy photons. The sensitivity is best for sources with
an extension around $0.5\deg$ and is worse for small and large sources.

Except for the hardest and largest sources, our sensitivity is almost
unchanged when using events between 100 \mev and 1 \gev.  This is also
demonstrated in figure~\ref{ts_ext_vs_energy} which shows the \tsext for
the SNR IC443 in 12 energy bins between 100 \mev and 100 \gev. For IC443,
which has a spectral index ~2.4, almost the entire increase in \ts comes
from energies above 1 \gev.  On the other hand, other systematic errors
become increasingly important at low energy. For our extension search,
we include only photons above 1 \gev.

Figure~\ref{diff_factor_sensitivity} shows how the LAT's
sensitivity varies with increasing background, typical of sources
at lower galactic latitude. We compute the sensitivity for spectral
index 2 sources using 10 and 100 times the
extragalactic background. The latter result approximately represents the intensity of
the galactic diffuse emission near the galactic center. 
The figure shows two plots, the first showing the sensitivity looking only between 
1 \gev and 100 \gev and the second showing the sensitivity looking only
between 10 \gev and 100 \gev.
For each plot, the $y$ axis quotes the flux in the selected energy range
which weaken the dependence of the sensitivity on spectral index.
This shows that we are less sensitive to extended sources in
a higher background region and that if the flux is constant,
we are more sensitive if the source is harder (has more high energy
photons).

@@

As was shown, we gain very little in sensitivity using photons with energies
below 1 \gev. On the other hand, the large PSF at low energy makes us
more susceptible to systematic errors arising from source confusion due
to multiple point sources and susceptible to incorrect modeling of the
galactic diffuse emission.  For that reason, we performed our search
using photons only between 1 \gev and 100 \gev.

We also performed a search for extended sources using photons only between
10 \gev and 100 \gev. Even though we are testing the same sources, this
approach is complimentary because there are regions in the galactic
plane where at lower energies the diffuse emission dominates and is
likely not perfectly modeled. On the other hand, at the highest energies
accessible by the LAT, the galactic diffuse emission will be smaller and
we might expect to detect significantly harder sources. This approach is
also beneficial because harder sources with nearby neighbors that are
softer may end up being incorrectly fit at lower energies but correctly
found looking only at the high energy.  This is also beneficial in the
galactic plane looking for possible extended sources near pulsars which
are very bright at lower energy but cutoff before 10 \gev.  This approach
of using only photons with energy above 10 \gev was successfully used
to detecting HESS J1825-137 with the LAT (\cite{fermi_hess_j1825}).

For each catalog source, we modified the spectral index of the source
to be a power-law. We did this for two reasons. The first is that for a
source to be significantly extended, we would expect it to be reasonably
hard and therefore be fit reasonably well by a power-law.  The second reason
is that since we begin our analysis at either 1 \gev or 10 \gev, we didn't
expect there to be significant curvature over this smaller energy range.
This method has the advantage that for the following analysis, we can
simply quote the fit power-law spectral index.

Finally, when analyzing our region, we automatically removed all other
catalog sources which were within 0.5\deg. This was done because of
a concern that truly extended sources might end up in the catalog as
multiple point sources. These spurious nearby catalogs sources might then
distort our fit enough that it no longer would show up as being extended.
Nevertheless, we investigated for each of our extended candidates if
if we seemed to be legitimately removing real background sources. And
we employed a dual localization procedure (described above) to test
all of our candidates for source confusion.

@@

We expect most extended sources to be inside the galactic plane
since extragalactic sources are typically too far away to be
resolvable. Unfortunately, the \gev emission in the galactic plane
is dominated by anisotropic galactic diffuse emission.  Modeling the
galactic diffuse emission and finding sources on top of this emission is
difficult~\ref{first_diffuse_paper}. This issue has been discussed in the
Fermi catalog~\ref{first_cat,second_cat}. Furthermore, the galactic plane is
very crowded and so it is often difficult to get a good model for the
region around a catalog source.  Because of this, finding a source with
$\tsext>16$ is not a sufficient test and we perform several crosschecks
to ensure our overall region is well modeled.

As an example of why this issue, we describe our test of P72Y3047 for
extension.  Figure~\ref{example_bad_fit} shows a smoothed counts map
using photons with energy above 1 \gev. In the map, the galactic and isotropic diffuse
emission subtracted.  A two year catalog source P72Y3047 was found at the
position $(l,b)=(37.79\deg,1.14\deg)$.  
We find that there appears to be overall some large scale residual in the
diffuse emission along the galactic plane. Furthermore, it appears that
we don't have a great model of the nearby catalog sources. As a result,
P72Y3047 fits to a very large extension of 1.37\deg and the result is
statistically significant with \tsext=51.6.  By looking at the residuals,
it is clear that this complicated region is not well fit even though
the source's extension is statistically significant. For this reason,
we have to perform a series of followup analysis for all of the sources
that are found with \tsext=16.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml

First, we took the spatial and spectral model found using \pointlike
and performed a crosscheck by refitting the spectral parameters using
\gtlike.  We used a square ROI $14.1\deg$ in diameter with $1/16\deg$
pixels and the same energy binning. We found that $1/16\deg$ bins were
fine enough for \gtlike to get an extension significances consistent
with \pointlike.  Using \gtlike, we found a second measure of \tsext and
we only considered a source to be significantly extended if \pointlike
and \gtlike both agreed.  For the new extended source candidates, with
table~\ref{alt_diff_model_results} shows a comparison of their the value
of \tsext found by \gtlike and \pointlike.  Everywhere else, the values
from \gtlike are used.

Additionally,
we generated for each extended source candidate residual
test statistic (\ts) maps.  The residual \ts map is a map of the \ts
value evaluated at each point by adding onto the map at that position
a new source of spectral index 2 and finding the increase in likelihood
when fitting its flux. \ts maps are especially useful as looking for
new sources or other residual features that are not included in the
model. They also have the advantageous feature that the absolute scale
can be used to clearly look for when the residual is significant.
The downside is they are biased to look for residual features
that look like the PSF and so are not as sensitive to diffuse residual
features. Furthermore, since by definition \ts>0, \ts maps can only show
positive fluctuations. Nevertheless, \ts maps were useful for
assessing the overall quality of the fit. Figure~\ref{res_tsmap}
shows an example residual \ts map for the extended source IC443.

For each candidate, we also generated counts slices. A counts slice
plots the sum of all counts in the longitudinal direction within 2\deg in
the latitudinal direction.  The same plot is also shown integrating in
the longitudinal direction.  On top of the plot is overlaid the model
predicted counts assuming the source has a disk spatial model, assuming
the source is a point source, and fitting just the diffuse emission in
the region.  An example counts slice is shows for the extended source
IC443 in figure~\ref{counts_slice}. Counts slices were especially useful
for looking at how well the overall diffuse emission model fits the region
near the sources.

For each candidate, we also made was a radial integral of the data. This
plot shows the sum of all counts within a given distance from the
source's center. The plot is binned uniformly in $\Delta \theta^2$ so
that an isotropic background will be flat in the plot. On top of this
plot is overlaid the model predicted counts assuming the source has a
disk spatial model, assuming the source is a point source, and fitting
just the diffuse emission in the region. An example radial integral
is shown for the extended source IC443 in figure~\ref{radial_profile}.
This plot is best for comparing the observed counts to the PSF.

Finally, for each source, we performed the dual localization method
described in section~\ref{dual_localization_method}.  We 
did so by replacing the source with two sources each containing
half of the original flux and then using \pointlike to
simultaneously fit the position and spectrum of these sources.
We then used \gtlike to refit the spectrum of these
sources using the best fit positions. For the
new extended sources found, we summarize the results of
this dual localization test and compare the values of \tsext
to \tsinc in table~\ref{dual_localization_results}.

@@

Since these sources are all included in the 2 year catalog, we applied
no special method to test these sources but used the same pipeline
that tested all other catalog sources for extension.  The results are
summarized in table~\ref{known_extended_sources}.  This table includes
the best fit position and extension. It also includes the source's
significance \ts and extension significance \tsext and the localization
error ellipse. The table also includes the best fit power-law spectral
model.

@@

\subsection{2FGL J1551.3-5333c}

% 1FGL J1554.0-5345c - P72Y2405         - 2FGL J1551.3-5333c
% Sources that had to be deleted
%   * P72Y2416 - 2FGL J1554.4-5317c
%   * P72Y2422 - Not a 2FGL source
